Experience-Biased Symbolic Planning

### 2014-05-11 AAAI_AuthorKit/paper.tex
SoCS Abstract-ready version

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
% AAAI format packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% Additional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Experience-Biased Symbolic Planning)
/Author (Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Experience-Biased Symbolic Planning}
\author{Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig\\
Carnegie Mellon University\\
5000 Forbes Ave,\\
Pittsburgh, PA 15213\\
}
\maketitle
\begin{abstract}
\begin{quote}
Weighted-heuristic A* search is a standard technique in currently state-of-the-art domain-independent symbolic planners. Experience graphs provide a means to augment existing heuristics with a bias toward using a select set of edges, typically taken from solutions to past queries. Thus, a planner can learn to bias its search toward well-understood regions and reuse parts of old plans. We show how to extend the HSP STRIPS planner with E-graphs at minimal cost. In doing so, we maintain the ability to configure path cost suboptimality guarantees, while in practice speeding up subsequent planning episodes across multiple domains.
\end{quote}
\end{abstract}

@@

\textbf{Definition+Theorem (needs TeX formatting):} a heuristic $h$ is called $\epsilon$-consistent if $h(S) = 0$ when $S$ is a goal and $h(S) \le \epsilon c(S,S') + h(S')$ for all $S,S'\in V$.
This property suffices to guarantee that each state is expanded only once by A*, and that the path it finds costs no more than $\epsilon$ times the optimal path. Consistency is synonymous with $1$-consistency.

@@

In the limit as $\epsilon^E \rightarrow 1$, following E-graph edges offers no advantage as $h$ is already an underestimate. Hence, using the triangle inequality to coalesce the sum,
\begin{eqnarray*}h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N h(S_{i-1},S_i) = \min_{N,S_N,\pi} h(S_0,S_N)
\\= \min_{S_N} h(S_0,S_N)\end{eqnarray*}

@@

In a preprocessing stage before the main search, we can apply Dijkstra's algorithm once in reverse with the costs $c'$ to compute the estimated distance-to-goal $h^E(S)$ from every $S\in V^E$. If the blackbox heuristic $h$ is cheap to compute, this preprocessing runs in $O(|V^E|^2)$ time, which is negligible for small E-graphs.

@@

$h^E$ is $\epsilon^E$-consistent [cite].

\section{Algorithm Discussion and Analysis}

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which is negligible for small E-graphs. A pseudocode implementation is listed in Algorithms \ref{alg:ComputeG} and \ref{alg:Search}.

@@

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG(S)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Do $\epsilon$-weighted A* from $I$ on full graph with heuristic $h^E$:
\FOR{$S \notin V^E$ when generated}
\STATE ComputeG(S)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG implements the dynamic programming computation of $g$-values. Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations. Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$. Thus, each state generated costs $O(\mathbf{A^2O + V^E})$ time, and the preprocessing cost is asymptotically equivalent to preemptively generating every state in $V^E$.

By comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it. The E-graph search maintains completeness and $\epsilon \epsilon^E$-suboptimality bounds on the path found. However, it makes no guarantee of improving the search time. Potential gains depend entirely on choosing a good small set of edges to call our E-graph. In our experiments we simply keep the previous solution path, but different choices remain a topic of investigation for future research.

>>>>

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
% AAAI format packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
% Additional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
% END Additional packages
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Experience-Biased Symbolic Planning)
/Author (Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Experience-Biased Symbolic Planning}
\author{Aram Ebtekar, Mike Phillips, Maxim Likhachev, Sven Koenig\\
Carnegie Mellon University\\
5000 Forbes Ave,\\
Pittsburgh, PA 15213\\
}
\maketitle
\begin{abstract}
\begin{quote}
Weighted A* search is the foundation for state-of-the-art domain-independent symbolic planners.
An important open challenge for these planners is to utilize experience with similar planning problems to speed up subsequent planning episodes.
Experience graphs have recently shown promise as a technique for plan reuse in the context of weighted A* searches in robotics, by biasing them towards a subgraph of the search space, typically chosen to consist of the edges used in previous plans.
In this paper, we demonstrate how to augment symbolic planners with
experience graphs, using the HSP STRIPS planner as a representative
example.
The resulting planners are able to trade off seamlessly
between plan quality and runtime and thus have advantages over past
plan-reuse techniques, such as planning by analogy.
We demonstrate
experimentally that our planner speeds up HSP by a factor of 2 or more on
many domains while achieving about the same plan quality.
\end{quote}
\end{abstract}

@@

\begin{defn} For $\epsilon\ge 1$, a heuristic $h$ is called \textbf{$\epsilon$-consistent} if $h(S) = 0$ when $S$ is a goal and $h(S) \le \epsilon c(S,S') + h(S')$ for all $S,S'\in V$. $h$ is \textbf{consistent} if it is $1$-consistent. \end{defn}
\begin{thm} Given an $\epsilon$-consistent heuristic and expanding no node more than once, A* is guaranteed to find a path costing no more than $\epsilon$ times the optimal path cost.  \end{thm}

@@

In the limit as $\epsilon^E \rightarrow 1$, following E-graph edges offers no advantage as $h$ is already an underestimate. Hence, using the triangle inequality to coalesce the sum,
\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N h(S_{i-1},S_i) = \min_{S_N} h(S_0,S_N)\]

@@

In a preprocessing stage before the main search, we apply Dijkstra's algorithm once in reverse with the costs $c'$ to compute the estimated distance-to-goal $h^E(S)$ from every $S\in V^E$. If the blackbox heuristic $h$ is cheap to compute, this preprocessing takes $O(|V^E|^2)$ time, which is negligible for small E-graphs.

@@

\begin{thm}$h^E$ is $\epsilon^E$-consistent. [cite]\end{thm}

This figure illustrates the idea. Dark edges have cost 3 and are in the E-graph. Light edges represent the jumps $h(S_i,S_j) = 2$ (with one exception: $h(S,G) = 5$), which we inflate by a factor $\epsilon^E=2$ to make 4 (or 10, respectively). Dijkstra's algorithm is used to compute the $h^E$ values shown alongside each node.

\section{Experience-Biased HSP}

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which is negligible for small E-graphs. A pseudocode implementation is listed in Algorithms \ref{alg:ComputeG} and \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while $\epsilon$ provides an additional goal-directed bias.

@@

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG(S)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated}
\STATE ComputeG(S)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG implements the dynamic programming computation of $g$-values. Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations. Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$. Thus, each state generated costs $O(\mathbf{A^2O + V^E})$ time, and the preprocessing cost is asymptotically equivalent to preemptively generating every state in $V^E$. For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it.

By $\epsilon\epsilon^E$-consistency of the weighted heuristic $\epsilon h^E$, the E-graph search is complete and the path it finds is guaranteed to be at most $\epsilon \epsilon^E$-suboptimal. However, it makes no guarantee of improving the search time. Potential gains depend entirely on choosing a good small set of edges to call our E-graph. In our experiments we simply keep the previous solution path; different choices remain a topic of investigation for future research.
